---
title: "Surface/bottom data summary"
author: "Abby Lewis"
date: "2024-08-08"
output: html_document
---

This file creates a summarized data frame with surface-water and bottom-water data for each lake/date, as well as maximum buoyancy frequency on that date. To calculate buoyancy frequency, we first interpolate all data to a 1m resolution.

## Table of contents

-   Step 1: Load packages and data
-   Step 2: Combine data
-   Step 3: Interpolate temperature
-   Step 4: Calculate buoyancy frequency
-   Step 5: Create df with surface and bottom measurements and buoyancy frequency

## Step 1: Load packages and data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pracma) # For interpolation function
library(tidyverse)
library(rLakeAnalyzer)
library(patchwork)

# Load data from EDI
data <- read_csv("https://pasta.lternet.edu/package/data/eml/edi/2220/6/ca2482ef3c43f053ae13cdf2bf5ba7a8?key=n9dYnNE5ReoA53SyzOCNgcPzLZA") %>%
  filter(Focal_site) # Not cleaned
cleaned_data <- read_csv("https://pasta.lternet.edu/package/data/eml/edi/2220/6/8553f79eeea9a6f722fc50d733ca9a74?key=n9dYnNE5ReoA53SyzOCNgcPzLZA") %>%
  filter(Focal_site) # Cleaned
temp <- read_csv("https://pasta.lternet.edu/package/data/eml/edi/2220/6/a101f4f31d3b2aa3dd81490892bd0b15?key=n9dYnNE5ReoA53SyzOCNgcPzLZA") %>%
  filter(Focal_site)
metadata <- read_delim("https://pasta.lternet.edu/package/data/eml/edi/2220/6/81f003749ad039837475571390b20190?key=n9dYnNE5ReoA53SyzOCNgcPzLZA") %>%
  filter(Focal_site)
```

## Step 2: Combine data

```{r combine}
# Match with metadata
profiles <- data %>%
  bind_rows(temp) %>%
  left_join(metadata) %>%
  rename(reported_max_depth = MaximumDepth_m) %>%
  dplyr::select(LakeID, Date, Depth_m, Temp_C, Source) %>%
  mutate(LakeID = paste(LakeID, Source, sep = "_"))

# summarize temps to depth and date, for lakes with high frequency measurements
profiles <- profiles %>%
  group_by(LakeID, Depth_m, Date, Source) %>%
  summarize(
    Temp_C = mean(Temp_C, na.rm = T)
  )
```

## Step 3: Interpolate temperature

```{r interpolate}
# Prep for interpolation
# We need at least three depths with water temp data
to_profiles_cleaning2 <- profiles %>%
  filter(!is.na(Temp_C) & !is.nan(Temp_C)) %>%
  arrange(LakeID, Date, Depth_m) %>%
  group_by(LakeID, Date) %>%
  mutate(somany = n_distinct(Depth_m)) %>%
  ungroup() %>%
  filter(
    somany > 2, # Can't do interpolation unless there are 3 depths
    !is.na(Date)
  ) %>%
  dplyr::select(-somany)

# Function to interpolate water temperature
interp_t <- function(lake, to_profiles_cleaning2) {
  alldates <- to_profiles_cleaning2 %>%
    dplyr::select(LakeID, Date, Depth_m, Temp_C) %>%
    filter(LakeID == lake) %>%
    dplyr::select(Date) %>%
    unique() %>%
    unlist() %>%
    as.vector()

  lake_data <- to_profiles_cleaning2 %>%
    filter(LakeID == lake)

  interpolatedvalues.do <- map_df(
    alldates,
    run_all_dates_t,
    lake_data, lake
  )

  return(interpolatedvalues.do)
}

# Function to run across all dates
run_all_dates_t <- function(alldates, lake_data, lake) {
  profile <- lake_data %>%
    filter(Date == alldates)

  datamatrix_ <- data.frame(
    depths = profile$Depth_m,
    values = profile$Temp_C
  ) %>%
    group_by(depths) %>%
    dplyr::summarize(valuemean = mean(values)) %>%
    ungroup() %>%
    as.matrix()

  max_depth <- max(profile$Depth_m, na.rm = T)
  interpolated.depths <- seq(
    from = min(profile$Depth_m), to = max(profile$Depth_m),
    by = ifelse(max_depth < 3, 0.25, 1)
  )

  if (max_depth <= 3) {
    interpolated.dos <- interp1(
      x = datamatrix_[, 1],
      y = datamatrix_[, 2],
      xi = interpolated.depths
    )
  } else {
    interpolated.dos <- pchip(
      xi = datamatrix_[, 1],
      yi = datamatrix_[, 2],
      x = interpolated.depths
    )
  }

  output <- data.frame(
    LakeID = lake,
    Date = alldates,
    Depth_m = interpolated.depths,
    Temp_C = interpolated.dos
  )

  return(output)
}

laken <- to_profiles_cleaning2$LakeID %>% unique()
interpolatedvalues.t <- map_df(laken, interp_t, to_profiles_cleaning2)
```

## Step 4: Calculate buoyancy frequency

```{r bf}
# Add shallow lakes back in (these wouldn't have been able to interpolate)
re_add <- profiles %>%
  filter(!is.na(Temp_C) & !is.nan(Temp_C)) %>%
  arrange(LakeID, Date, Depth_m) %>%
  group_by(LakeID, Date) %>%
  filter(n_distinct(Depth_m) == 2) %>%
  left_join(metadata %>%
    mutate(LakeID = paste(LakeID, Source, sep = "_")) %>%
    select(LakeID, MaximumDepth_m)) %>%
  filter(MaximumDepth_m < 3) %>%
  select(-Source, -MaximumDepth_m)

database.interp <- interpolatedvalues.t %>%
  mutate(Date = as.Date(Date, origin = "1970-01-01")) %>%
  bind_rows(re_add) %>%
  group_by(LakeID, Date) %>%
  arrange(Depth_m, .by_group = T)

buoyancy_freq <- database.interp %>%
  group_by(LakeID, Date) %>%
  filter(
    !is.na(Temp_C),
    !is.na(Depth_m),
    n() >= 3
  ) %>% # Need 3 measurements for thermo depth
  dplyr::summarize(buoyancy_frequency = max(rLakeAnalyzer::buoyancy.freq(Temp_C, Depth_m))) %>%
  ungroup() %>%
  distinct() %>%
  separate(LakeID, into = c("LakeID", "Source"), sep = "_") %>%
  left_join(metadata %>% dplyr::select(LakeID, LakeName, Source), by = c("LakeID", "Source"))
```

## Step 5: Create df with surface and bottom measurements and buoyancy frequency

```{r comb}
layer_sum_not_mean <- cleaned_data %>%
  select(LakeID, Source, Date, Depth_m, CO2_umolL, CH4_umolL, DO_mgL, Temp_C, TP_ugL) %>%
  pivot_longer(CO2_umolL:TP_ugL) %>%
  group_by(LakeID, Source, Date, name) %>%
  summarize(
    surf_depth = min(Depth_m, na.rm = T),
    surf_value = value[Depth_m == surf_depth],
    bot_depth = max(Depth_m, na.rm = T),
    bot_value = mean(value[Depth_m == bot_depth], na.rm = T)
  ) %>%
  pivot_longer(
    cols = c(surf_depth, surf_value, bot_depth, bot_value),
    names_to = c("Layer", ".value"),
    names_pattern = "(surf|bot)_(.*)"
  ) %>%
  pivot_wider(
    names_from = name,
    values_from = value
  ) %>%
  rename(Depth_m = depth) %>%
  left_join(metadata) %>%
  filter(Layer == "surf" & (Depth_m < 1 / 3 * MaximumDepth_m) |
    Layer == "bot" & (Depth_m > 2 / 3 * MaximumDepth_m)) %>%
  select(LakeID:Temp_C) %>%
  left_join(buoyancy_freq %>% select(LakeID, Source, Date, buoyancy_frequency) %>% distinct())

write_csv(layer_sum_not_mean, "../CompiledData/layer_sum.csv")

# Stats for MS
meta_corrected <- read_csv("../CompiledData/metadata_correct_depths.csv") 

#' Surface measurements
#' Mean depth - 0.77 m, 7% of maximum depth
layer_sum <- layer_sum_not_mean %>% left_join(meta_corrected)
layer_sum %>% filter(Layer == "surf") %>% pull(Depth_m) %>% mean()
layer_sum %>% filter(Layer == "surf") %>% mutate(ratio = Depth_m/MaximumDepth_m) %>% pull(ratio) %>% mean(na.rm = T)

a <- layer_sum %>% filter(Layer == "surf") %>% mutate(ratio = Depth_m/MaximumDepth_m) %>%
  ggplot(aes(x = ratio*100)) +
  geom_histogram(binwidth = 2) +
  labs(x = "Perecent of max depth", y = "Count of 'surface' measurements") +
  theme_bw()

#' Bottom measurements
#' Mean depth - 18m, 90% of maximum depth
layer_sum %>% filter(Layer == "bot") %>% pull(Depth_m) %>% mean()
layer_sum %>% filter(Layer == "bot") %>% mutate(ratio = Depth_m/MaximumDepth_m) %>% pull(ratio) %>% mean(na.rm = T)

b <- layer_sum %>% filter(Layer == "bot") %>% mutate(ratio = Depth_m/MaximumDepth_m) %>%
  ggplot(aes(x = ratio*100)) +
  geom_histogram(binwidth = 2) +
  labs(x = "Perecent of max depth", y = "Count of 'bottom' measurements") +
  scale_x_continuous(limits = c(60,100)) +
  theme_bw()

a+b
ggsave("../Figures/R2.png", width = 6, height = 3.5)

test2 <- layer_sum %>% filter(Layer == "bot") %>% mutate(ratio = Depth_m/MaximumDepth_m) %>%
  filter(ratio < 0.66)

#' Mean thermocline depth is 33% of the water column

layer_sum_stats <- layer_sum_not_mean %>%
  left_join(metadata)

median(layer_sum_stats$Depth_m[layer_sum_stats$Layer == "surf"])
median(
  layer_sum_stats$Depth_m[layer_sum_stats$Layer == "bot"] /
    layer_sum_stats$MaximumDepth_m[layer_sum_stats$Layer == "bot"],
  na.rm = T
)
```

Correcting depth in metadata

```{r}
true_max_depths <- layer_sum_not_mean %>% left_join(metadata) %>%
  group_by(LakeID) %>% summarize(
        true_max_depth = ifelse(max(Depth_m, na.rm = T) > max(MaximumDepth_m), 
                                max(Depth_m, na.rm = T), max(MaximumDepth_m))
  )

metadata_correct_max_depth <- metadata %>% left_join(true_max_depths) %>%
  mutate(MaximumDepth_m = ifelse(true_max_depth > MaximumDepth_m, 
                                 true_max_depth, MaximumDepth_m)) %>%
  select(-true_max_depth)

write_csv(metadata_correct_max_depth, "../CompiledData/metadata_correct_depths.csv")
```

