---
title: "Concentration boxplots"
format: html
editor: visual
---

This file creates summary boxplots for the Supplemental Information.

## Table of contents

-   Step 1: Load data and packages
-   Step 2: Plot concentration by mixing regime
-   Step 3: Plot concentration and ratio by lake type
-   Step 4: Simpler summary plot for presentations

## Step 1: Load data and packages

```{r setup}
# Load packages
library(tidyverse)

# Load data from EDI
data <- read_csv("https://pasta.lternet.edu/package/data/eml/edi/2220/5/8553f79eeea9a6f722fc50d733ca9a74") %>%
  filter(Focal_site) # Cleaned
metadata <- read_delim("https://pasta.lternet.edu/package/data/eml/edi/2220/5/81f003749ad039837475571390b20190") %>%
  filter(Focal_site)
layer_sum <- read_csv("../CompiledData/layer_sum.csv")
```

## Step 2: Plot concentration by mixing regime

```{r mix_troph}
mix_troph <- layer_sum %>%
  left_join(metadata) %>%
  mutate(
    TrophicStatus = ifelse(is.na(TrophicStatus) | TrophicStatus == "other",
      "not reported",
      TrophicStatus
    ),
    TrophicStatus = factor(TrophicStatus,
      levels = c(
        "oligotrophic",
        "mesotrophic",
        "eutrophic",
        "not reported"
      )
    ),
    StratificationRegime = factor(StratificationRegime,
      levels = c(
        "meromictic or amictic",
        "monomictic",
        "dimictic",
        "dimictic to polymictic",
        "polymictic"
      ),
      labels = c(
        "meromictic\nor amictic",
        "monomictic",
        "dimictic",
        "dimictic to\npolymictic",
        "polymictic"
      )
    )
  ) %>%
  pivot_longer(c(CO2_umolL, CH4_umolL), names_to = "gas") %>%
  mutate(
    gas = case_match(
      gas,
      "CO2_umolL" ~ "CO[2]",
      "CH4_umolL" ~ "CH[4]"
    ),
    Layer = factor(Layer,
      levels = c("surf", "bot"),
      labels = c("Surface", "Bottom")
    )
  ) %>%
  group_by(gas, Layer, LakeID, TrophicStatus, StratificationRegime) %>%
  summarize(max = max(value, na.rm = T)) %>%
  filter(!is.na(StratificationRegime), !is.na(max)) %>%
  ggplot(aes(x = StratificationRegime, y = max)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2, aes(color = TrophicStatus), alpha = 0.2) +
  scale_y_log10() +
  ylab(expression("Concentration (" * mu * "M)")) +
  theme_bw() +
  facet_grid(gas ~ Layer, scales = "free", labeller = label_parsed) +
  scale_color_manual(values = c("navy", "turquoise3", "green4", "grey85")) +
  egg::theme_article() +
  theme(
    axis.title.x = element_blank(),
    legend.title = element_blank(),
    panel.grid.major = element_line(color = "grey90", size = 0.5),
    panel.grid.minor = element_line(color = "grey95", size = 0.25),
    axis.text.x = element_text(
      angle = 40,
      vjust = 1,
      hjust = 1,
      lineheight = 0.6
    )
  )

ggsave("../Figures/Boxplot_mixing_trophic.jpeg",
  plot = mix_troph,
  width = 6, height = 4, units = "in", dpi = 300
)
```

## Step 3: Plot concentration and ratio by lake type

```{r type_troph}
# Calculate bottom:surface ratios
fract <- layer_sum %>%
  pivot_longer(c(CO2_umolL, CH4_umolL), names_to = "gas") %>%
  group_by(gas, Layer, LakeID, Source) %>%
  summarize(max = max(value, na.rm = T)) %>%
  filter(!is.na(max)) %>%
  pivot_wider(names_from = Layer, values_from = max) %>%
  filter(
    !is.na(bot),
    !is.na(surf)
  ) %>%
  mutate(fract = bot / surf) %>%
  select(fract, gas, LakeID, Source)

data_for_box2 <- layer_sum %>%
  left_join(metadata) %>%
  mutate(
    TrophicStatus = ifelse(is.na(TrophicStatus) | TrophicStatus == "other",
      "not reported",
      TrophicStatus
    ),
    TrophicStatus = factor(TrophicStatus,
      levels = c(
        "oligotrophic",
        "mesotrophic",
        "eutrophic",
        "not reported"
      ),
      labels = c(
        "Oligotrophic",
        "Mesotrophic",
        "Eutrophic",
        "Not reported"
      )
    ),
    LakeType = factor(LakeType,
      levels = c(
        "reservoir",
        "lake",
        "small lake",
        "shallow lake or wetland",
        "pond"
      ),
      labels = c(
        "Reservoir",
        "Lake",
        "Small lake",
        "Shallow lake",
        "Pond"
      )
    )
  ) %>%
  pivot_longer(c(CO2_umolL, CH4_umolL), names_to = "gas") %>%
  mutate(gas = factor(gas,
    levels = c("CH4_umolL", "CO2_umolL"),
    labels = c("CH[4]", "CO[2]")
  )) %>%
  group_by(gas, Layer, LakeID, TrophicStatus, LakeType) %>%
  summarize(max = max(value, na.rm = T)) %>%
  filter(!is.na(LakeType), !is.na(max), is.finite(max)) %>%
  pivot_wider(names_from = Layer, values_from = max) %>%
  mutate(
    surf = ifelse(surf == 0, 0.001, surf),
    fract = bot / surf
  ) %>%
  pivot_longer(c(surf, bot, fract), names_to = "Layer", values_to = "max") %>%
  mutate(Layer = factor(Layer,
    levels = c("surf", "bot", "fract"),
    labels = c(
      "Surface", "Bottom",
      "Bottom:surface"
    )
  ))

stats <- data_for_box2 %>%
  filter(
    !is.na(max),
    !is.infinite(max)
  ) %>%
  group_by(Layer, gas) %>%
  summarize(
    kruskal_p = kruskal.test(max ~ LakeType)$p.value,
    p_vals = dunn.test::dunn.test(max,
      LakeType,
      method = "Bonferroni"
    )$P.adjusted,
    names = gsub(" ", "", dunn.test::dunn.test(max,
      LakeType,
      method = "Bonferroni"
    )$comparisons)
  ) %>%
  reframe(rcompanion::cldList(
    comparison = names,
    p.value = p_vals
  )) %>%
  mutate(LakeType = case_match(Group,
    "Shallowlake" ~ "Shallow lake",
    "Smalllake" ~ "Small lake",
    .default = Group
  ))

type_troph <- data_for_box2 %>%
  ggplot(aes(x = LakeType, y = data_for_box2$max)) +
  geom_hline(aes(yintercept = y),
    data = data.frame(
      Layer = "fract",
      y = 1
    ) %>%
      mutate(Layer = factor(Layer,
        levels = c("surf", "bot", "fract"),
        labels = c(
          "Surface", "Bottom",
          "Bottom:surface"
        )
      ))
  ) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2, aes(color = TrophicStatus), alpha = 0.2) +
  geom_text(data = stats, aes(label = Letter, y = Inf), vjust = 1.2) +
  scale_y_log10(
    labels = scales::label_comma(drop0trailing = T),
    expand = expansion(mult = c(0, .15))
  ) +
  ylab(expression("Concentration (" * mu * "M)")) +
  theme_bw() +
  ggh4x::facet_grid2(gas ~ Layer,
    scales = "free", independent = "y",
    labeller = label_parsed,
    switch = "y"
  ) +
  scale_color_manual(
    values = c("navy", "turquoise3", "green4", "grey70"),
    name = "Trophic state"
  ) +
  egg::theme_article() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(
      angle = 40,
      vjust = 1,
      hjust = 1,
      lineheight = 0.6
    ),
    panel.grid.major = element_line(color = "grey90", size = 0.5),
    panel.grid.minor = element_line(color = "grey95", size = 0.25),
    strip.placement = "outside",
    legend.position = "bottom"
  )

ggsave("../Figures/Boxplot_type_trophic.jpeg",
  plot = type_troph,
  dpi = 300, width = 6.5, height = 4.5
)
```

## Step 4: Simpler summary plot for presentations

```{r simple_sum}
summary_talk <- layer_sum %>%
  filter(Layer == "bot") %>%
  left_join(metadata) %>%
  mutate(
    TrophicStatus = ifelse(is.na(TrophicStatus) | TrophicStatus == "other",
      "not reported",
      TrophicStatus
    ),
    TrophicStatus = factor(TrophicStatus,
      levels = c(
        "oligotrophic",
        "mesotrophic",
        "eutrophic",
        "not reported"
      )
    ),
    StratificationRegime = factor(StratificationRegime,
      levels = c(
        "meromictic or amictic",
        "monomictic",
        "dimictic",
        "dimictic to polymictic",
        "polymictic"
      ),
      labels = c(
        "meromictic\nor amictic",
        "monomictic",
        "dimictic",
        "dimictic to\npolymictic",
        "polymictic"
      )
    )
  ) %>%
  group_by(LakeID, TrophicStatus, StratificationRegime) %>%
  summarize(max = max(CH4_umolL, na.rm = T)) %>%
  filter(!is.na(StratificationRegime), !is.na(max)) %>%
  ggplot(aes(x = StratificationRegime, y = max)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2, aes(color = TrophicStatus)) +
  scale_y_log10() +
  ylab(bquote("Deep-water" ~ CH[4] ~ "(µmol/L)")) +
  theme_bw() +
  scale_color_manual(values = c("navy", "turquoise3", "green4", "grey85")) +
  theme(
    axis.title.x = element_blank(),
    legend.title = element_blank()
  )

jpeg("../Figures/Hypo. CH4 vs Strat.jpeg", width = 4, height = 3.5, units = "in", res = 300)
summary_talk +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "bottom",
    legend.direction = "vertical"
  ) +
  guides(color = guide_legend(ncol = 3))
dev.off()
```
