---
title: "Random forest analysis"
format: html
editor: visual
---

This file runs random forest regressions to characterize the drivers of surface- and bottom-water GHG concentrations and saturation, as well as the ratio of bottom to surface concentrations.

## Table of contents

-   Step 1: Load data and packages
-   Step 2: Format data for regressions
-   Step 3: Plot summary stats about drivers
-   Step 4: SPLM analysis
-   Step 5: Permutation
-   Step 6: Simple/complex comparison
-   Step 7: Ratio analysis
-   Step 8: Percent saturation

## Step 1: Load data and packages

```{r setup, include = FALSE}
# Load packages
library(tidyverse)
library(randomForest)
library(ggh4x)
library(rLakeAnalyzer)
library(marelac)
source("R/equilibrium_calcs.R")
source("R/generate_plot_splmRF.R")
source("R/generate_plot_permutation.R")
source("R/generate_plot_ratio_permutation.R")
source("R/comp_two_rf_permutation.R")

# Load data
metadata <- read_delim("https://pasta.lternet.edu/package/data/eml/edi/2220/5/81f003749ad039837475571390b20190") %>%
  filter(Focal_site)
layer_sum <- read_csv("../CompiledData/layer_sum.csv")

# Create dictionary to use throughout
var_names <- c(
  "DO_mgL" = "DO~(mg~L^-1)",
  "SurfaceArea_km2" = "SA~(km^2)",
  "MaximumDepth_m" = "Max~depth~(m)",
  "MeanDepth_m" = "Mean~depth~(m)",
  "TP_ugL_mean" = "TP~(µg~L^-1)",
  "Temp_C" = "Temp~(ºC)",
  "Osgood" = "Osgood",
  "Absolute_latitude" = "Abs~lat~(DD)",
  "buoyancy_frequency" = "Buoy~freq~(s^-2)",
  "daylength" = "Day~length~(h)",
  "dens_dif" = "dens_dif",
  "Latitude" = "Latitude",
  "TrophicStatus" = "Trophic\nstatus",
  "StratificationRegime" = "Stratification\nregime",
  "bot_DO_mgL" = "Bot~DO~(mg~L^-1)",
  "bot_buoyancy_frequency" = "Buoy~freq~(s^-2)",
  "yday" = "Day~of~year",
  "LakeID" = "Lake~random~effect"
)

log_vars <- c(
  "SurfaceArea_km2", "MeanDepth_m",
  "MaximumDepth_m", "TP_ugL_mean",
  "bot_buoyancy_frequency", "Osgood",
  "buoyancy_frequency"
)

# Set seed to produce consistent RF results
set.seed(47)
```

## Step 2: Format data for regressions

```{r format}
layer_plot_data <- layer_sum %>%
  mutate(density = water.density(Temp_C)) %>%
  group_by(LakeID, Date) %>%
  mutate(dens_dif = ifelse(length(density) == 2,
    abs(density[Layer == "bot"] -
      density[Layer == "surf"]),
    NA
  )) %>%
  pivot_longer(cols = CO2_umolL:CH4_umolL) %>%
  group_by(LakeID, name, Layer, Date) %>%
  filter(!is.na(value)) %>%
  mutate(
    value = mean(value, na.rm = T),
    yday = yday(Date),
    year = year(Date)
  ) %>%
  group_by(name, Layer) %>%
  mutate(name = ifelse(name == "CO2_umolL",
    "CO[2]",
    "CH[4]"
  ))

metadata_by_lake <- metadata %>%
  group_by(LakeID) %>%
  mutate(
    across(
      c(TrophicStatus, StratificationRegime),
      \(x) ifelse(is.na(x), unique(x[!is.na(x)]), x)
    ),
    Osgood = MeanDepth_m / sqrt(SurfaceArea_km2)
  ) %>%
  summarize(
    across(
      c(
        MaximumDepth_m,
        SurfaceArea_km2,
        Chla_ugL_mean,
        TP_ugL_mean,
        DOC_mgL_mean,
        MeanDepth_m,
        Osgood,
        Latitude,
        Longitude
      ),
      \(x) mean(x, na.rm = T)
    ),
    across(
      c(
        TrophicStatus,
        StratificationRegime,
        LakeType
      ),
      \(x) paste(unique(x), collapse = " AND ")
    )
  )

all <- layer_plot_data %>%
  left_join(metadata_by_lake) %>%
  mutate(
    TrophicStatus = ifelse(grepl(" AND ", TrophicStatus),
      NA,
      TrophicStatus
    ),
    StratificationRegime = ifelse(grepl(" AND ", StratificationRegime),
      NA,
      StratificationRegime
    ),
    value = ifelse(value <= 0, 0.001, value),
    value = log(value),
    TrophicStatus = factor(TrophicStatus,
      levels = c(
        "oligotrophic",
        "mesotrophic",
        "eutrophic"
      )
    ),
    StratificationRegime = factor(StratificationRegime,
      levels = c(
        "meromictic or amictic",
        "monomictic",
        "dimictic",
        "dimictic to polymictic",
        "polymictic"
      ),
      labels = c(
        "meromictic\nor amictic",
        "monomictic",
        "dimictic",
        "dimictic to\npolymictic",
        "polymictic"
      )
    ),
    Absolute_latitude = abs(Latitude),
    daylength = geosphere::daylength(Latitude, yday),
    yday = ifelse(Latitude < 0,
      ifelse(yday + 365 / 2 > 365,
        yday - 365 / 2,
        yday + 365 / 2
      ),
      yday
    )
  )

# The order of these controls the order of the drivers in importance/partial plots
vars <- c(
  "DO_mgL",
  "SurfaceArea_km2",
  "Osgood",
  "TP_ugL_mean",
  "Temp_C",
  "Absolute_latitude",
  "Latitude", # Not used as a driver in permutation approach
  "Longitude", # Not used as a driver in permutation approach
  "LakeID" # Not used as a driver in permutation approach
)

log_select <- log_vars[log_vars %in% vars]
```

## Step 3: Plot summary stats about drivers

```{r}
for_cor <- all %>%
  filter(if_all(all_of(log_select), \(x) x > 0)) %>%
  ungroup() %>%
  mutate(
    across(all_of(log_select), log),
    mutate(across(where(is.character), as.factor))
  ) %>%
  select(all_of(c("value", vars, "name", "Layer", "LakeID", "Date", "Source"))) %>%
  na.omit() %>%
  ungroup() %>%
  pivot_wider(names_from = Layer, values_from = c(value, DO_mgL, Temp_C)) %>%
  pivot_wider(names_from = name, values_from = c(value_surf, value_bot)) %>%
  select(
    where(is.numeric),
    -Longitude,
    -Latitude
  )

dens_plots <- for_cor %>%
  pivot_longer(everything()) %>%
  mutate(name = factor(name,
    levels = c(
      "DO_mgL_bot", "DO_mgL_surf", "SurfaceArea_km2", "Osgood",
      "TP_ugL_mean", "Temp_C_bot", "Temp_C_surf",
      "Absolute_latitude",
      "value_surf_CH[4]", "value_bot_CH[4]", "value_surf_CO[2]",
      "value_bot_CO[2]"
    ),
    labels = c(
      "Bottom~DO~(mg~L^-1)", "Surface~DO~(mg~L^-1)", "SA~(log~km^2)",
      "Osgood~index~(log)",
      "TP~(log~µg~L^-1)", "Bottom~temp~(ºC)", "Surface~temp~(ºC)",
      "Abs.~latitude",
      "Surface~CH[4]~(log~µM)", " Bottom~CH[4]~(log~µM)",
      "Surface~CO[2]~(log~µM)",
      "Bottom~CO[2]~(log~µM)"
    )
  )) %>%
  ggplot(aes(x = value)) +
  geom_density() +
  facet_wrap(~name, scales = "free", labeller = "label_parsed") +
  egg::theme_article()

ggsave("../Figures/dens_plots.png", dens_plots, width = 7, height = 4.5, dpi = 300)

for_corrplot <- for_cor %>%
  mutate(row = row_number()) %>%
  pivot_longer(-row) %>%
  filter(!grepl("value", name)) %>%
  mutate(name = factor(name,
    levels = c(
      "DO_mgL_bot", "DO_mgL_surf", "SurfaceArea_km2", "Osgood",
      "TP_ugL_mean", "Temp_C_bot", "Temp_C_surf",
      "Absolute_latitude"
    ),
    labels = c(
      ":Bottom~DO~(mg~L^-1)", ":Surface~DO~(mg~L^-1)", ":SA~(log~km^2)",
      ":Osgood~index~(log)",
      ":TP~(log~µg~L^-1)", ":Bottom~temp~(ºC)", ":Surface~temp~(ºC)",
      ":Abs.~latitude"
    )
  )) %>%
  pivot_wider() %>%
  select(-row) %>%
  na.omit()

col <- colorRampPalette(rev(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA")))
M <- cor(for_corrplot, method = "spearman")

jpeg("../Figures/corrplot.png", width = 6.5, height = 6.5, res = 300, units = "in")
corrplot::corrplot(M,
  method = "color", col = col(200),
  type = "upper", order = "hclust",
  addCoef.col = "black", # Add coefficient of correlation
  tl.col = "black", tl.srt = 45, # Text label color and rotation
  # Combine with significance
  sig.level = 0.01, insig = "blank",
  # hide correlation coefficient on the principal diagonal
  diag = FALSE,
  mar = c(0, 0, 0, 0)
)
dev.off()
```

## Step 4: SPLM analysis

```{r splm}
# Run regressions
co2_plot <- generate_plot_splmRF(
  all = all, vars = vars, var_names = var_names,
  gas_name = "CO[2]", log_vars = log_vars
)
co2_plot
ggsave(co2_plot,
  dpi = 300,
  filename = "../Figures/CO2 drivers - surf bot - splmRF.jpeg",
  width = 7, height = 6
)

ch4_plot <- generate_plot_splmRF(all, vars, var_names, "CH[4]", log_vars)
ch4_plot
ggsave(ch4_plot,
  dpi = 300,
  filename = "../Figures/CH4 drivers - surf bot - splmRF.jpeg",
  width = 7, height = 6
)
```

## Step 5: Permutation

```{r permutation}
# Run regressions
co2_plot <- generate_plot_permutation(
  all = all, vars = vars,
  var_names = var_names, gas_name = "CO[2]",
  log_vars = log_vars, reps = 10
)
co2_plot
ggsave(co2_plot,
  dpi = 300,
  filename = "../Figures/CO2 drivers - surf bot - permutation.jpeg",
  width = 7, height = 6
)

ch4_plot <- generate_plot_permutation(
  all = all, vars = vars,
  var_names = var_names, gas_name = "CH[4]",
  log_vars = log_vars, reps = 10
)
ch4_plot
ggsave(ch4_plot,
  dpi = 300,
  filename = "../Figures/CH4 drivers - surf bot - permutation.jpeg",
  width = 7, height = 6
)

# Consistent
all_consistent <- all %>%
  ungroup() %>%
  select(all_of(c("value", "name", "Layer", "LakeID", vars))) %>%
  drop_na() %>%
  group_by(LakeID) %>%
  filter(
    sum(Layer == "bot" & name == "CO[2]") >= 1,
    sum(Layer == "bot" & name == "CH[4]") >= 1,
    sum(Layer == "surf" & name == "CO[2]") >= 1,
    sum(Layer == "surf" & name == "CH[4]") >= 1
  )

co2_plot <- generate_plot_permutation(
  all = all_consistent, vars = vars,
  var_names = var_names, gas_name = "CO[2]",
  log_vars = log_vars, reps = 10
)
co2_plot
ggsave(co2_plot,
  dpi = 300,
  filename = "../Figures/CO2 drivers - surf bot - permutation - consistent.jpeg",
  width = 7, height = 6
)

ch4_plot <- generate_plot_permutation(
  all = all_consistent, vars = vars,
  var_names = var_names, gas_name = "CH[4]",
  log_vars = log_vars, reps = 10
)
ch4_plot
ggsave(ch4_plot,
  dpi = 300,
  filename = "../Figures/CH4 drivers - surf bot - permutation - consistent.jpeg",
  width = 7, height = 6
)

# Lakes with one measurement
all %>%
  filter(
    if_all(all_of(log_select), \(x) x > 0)
  ) %>%
  ungroup() %>%
  select(all_of(c("value", vars, "LakeID", "name", "Layer", "Date"))) %>%
  na.omit() %>%
  group_by(LakeID, name, Layer) %>%
  summarize(
    n = length(unique(Date))
  ) %>%
  group_by(name, Layer) %>%
  summarize(
    only_1 = sum(n == 1),
    median = median(n)
  )
```

## Step 6: Simple/complex comparison

```{r simple_complex}
vars_simple <- c(
  "SurfaceArea_km2",
  "Osgood",
  "TrophicStatus",
  "Absolute_latitude"
)

s <- comp_two_rf_permutation(all, vars_simple, vars,
  log_vars,
  reps = 10
)

ggsave(s, dpi = 300, filename = "../Figures/Model comparison - rmse comb imp - permutation.jpeg", width = 6, height = 5)
```

## Step 7: Ratio analysis

```{r ratio}
epi_hypo <- layer_sum %>%
  select(-Depth_m) %>%
  pivot_longer(
    cols = CH4_umolL:buoyancy_frequency,
    names_to = "variable", values_to = "value"
  ) %>%
  pivot_wider(names_from = c(Layer, variable), values_from = value) %>%
  mutate(
    ratio_co2 = bot_CO2_umolL / surf_CO2_umolL,
    ratio_ch4 = bot_CH4_umolL / surf_CH4_umolL
  ) %>%
  left_join(metadata_by_lake) %>%
  mutate(Absolute_latitude = abs(Latitude))

epi_hypo_co2 <- epi_hypo %>%
  group_by(LakeID, Date) %>%
  filter(is.finite(ratio_co2)) %>%
  mutate(
    name = "ratio_co2",
    value = ifelse(ratio_co2 <= 0, 0.001, ratio_co2),
    value = log(value)
  )

epi_hypo_ch4 <- epi_hypo %>%
  group_by(LakeID, Date) %>%
  filter(is.finite(ratio_ch4)) %>%
  mutate(name = "ratio_ch4") %>%
  mutate(
    value = ifelse(ratio_ch4 <= 0, 0.001, ratio_ch4),
    value = log(value)
  )

vars <- c(
  "bot_DO_mgL",
  "SurfaceArea_km2",
  "Osgood",
  "TP_ugL_mean",
  "bot_buoyancy_frequency",
  "Absolute_latitude"
)

# Permutation
comb_plot <- generate_plot_ratio_permutation(
  all_co2 = epi_hypo_co2,
  all_ch4 = epi_hypo_ch4,
  vars = vars, var_names = var_names,
  log_vars = log_vars,
  reps = 10
)
comb_plot
ggsave(comb_plot,
  dpi = 300,
  filename = "../Figures/Ratio drivers - permutation.jpeg",
  width = 6, height = 5
)
```

## Step 8: Percent saturation

```{r pSat}
all_pSat <- all %>%
  mutate(
    month = month(as.Date(paste0(year, "-01-01")) + days(round(yday, 0) - 1)),
    value = case_when(
      name == "CO[2]" ~ (exp(value) / calc_CO2_equil(Temp_C, year, month)) * 100,
      name == "CH[4]" ~ (exp(value) / calc_CH4_equil(Temp_C, year, month)) * 100
    ),
    value = log(value)
  ) %>%
  filter(!is.na(value))

vars <- c(
  "DO_mgL",
  "SurfaceArea_km2",
  "Osgood",
  "TP_ugL_mean",
  "Temp_C",
  "Absolute_latitude"
)

co2_plot <- generate_plot_permutation(
  all = all_pSat, vars = vars,
  var_names = var_names, pSat = T,
  gas_name = "CO[2]", log_vars = log_vars,
  reps = 10
)
co2_plot
ggsave(co2_plot,
  dpi = 300,
  filename = "../Figures/CO2 drivers - surf bot - pSat - permutation.jpeg",
  width = 7, height = 6
)

ch4_plot <- generate_plot_permutation(
  all = all_pSat, vars = vars,
  var_names = var_names, pSat = T,
  gas_name = "CH[4]", log_vars = log_vars,
  reps = 10
)
ch4_plot
ggsave(ch4_plot,
  dpi = 300,
  filename = "../Figures/CH4 drivers - surf bot - pSat - permutation.jpeg",
  width = 7, height = 6
)
```
