---
title: "Ecosystems"
format: html
editor: visual
---

This file compares gas concentrations across ecosystems.

## Table of Contents

-   Step 1: Load packages
-   Step 2: Format our data
-   Step 3: Streams and rivers
-   Step 4: Oceans
-   Step 5: Estuaries
-   Step 6: Wetlands
-   Step 7: Combine and plot
-   Step 8: statistics

## Step 1: Load packages

```{r packages}
library(tidyverse)
source("R/make_ecosystem_plot.R")
```

## Step 2: Format our data

```{r ours}
# Load data
layer_sum <- read_csv("../CompiledData/layer_sum.csv")

layer_means <- layer_sum %>%
  pivot_longer(cols = CO2_umolL:CH4_umolL) %>%
  group_by(LakeID, name, Layer, Date) %>%
  filter(!is.na(value)) %>%
  mutate(
    value = mean(value, na.rm = T),
    yday = yday(Date),
    year = year(Date),
    
  ) %>%
  group_by(LakeID, name, Layer) %>%
  summarize(
    value = mean(value, na.rm = T)
  ) %>%
  rename(Ecosystem = Layer) %>%
  mutate(Ecosystem = ifelse(Ecosystem == "surf",
                            "Lake\nsurface",
                            "Lake\nbottom"))
```

## Step 3: Streams and rivers

```{r streams_and_rivers}
#https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-ntl.420.3
#https://essd.copernicus.org/articles/15/2879/2023/essd-15-2879-2023.html
streams <- read_csv("https://pasta.lternet.edu/package/data/eml/knb-lter-ntl/420/3/ba3e270bcab8ace5d157c995e4b791e4?key=n9dYnNE5ReoA53SyzOCNgcPzLZA")

stream_sum = streams %>%
  filter(CH4mean >= 0) %>%
  group_by(Source_ID, Site_ID) %>%
  summarize(value = mean(CH4mean, na.rm = T)) %>%
  mutate(name = "CH4_umolL",
         Ecosystem = "Streams and rivers",
         Site_ID = as.character(Site_ID))

#https://www.nature.com/articles/s41597-022-01915-0#Sec6
#https://figshare.com/articles/dataset/CDFLOW_-_A_30-year_dataset_on_CO2_in_flowing_freshwaters_in_the_United_States/19787326

flow <- read_csv("../CompiledData/CDFLOW.csv") %>%
  filter(`CO2.mg/l` >= 0) %>%
  group_by(COMID) %>%
  summarize(name = "CO2_umolL",
            value = mean(`CO2.mg/l`, na.rm = T) * 1000 / 44.01) %>%
  mutate(Ecosystem = "Streams and rivers")

#https://www.nature.com/articles/s44221-025-00522-8
tropical_rivers <- read_csv("https://raw.githubusercontent.com/ClementDuvert/GHG.flux.tropics/refs/heads/main/Data/dataset_rivers_reduced.csv")

tropical_rivers_sum <- tropical_rivers %>%
  select(CO2_uM, CH4_uM) %>%
  rename(CO2_umolL = CO2_uM,
         CH4_umolL = CH4_uM) %>%
  pivot_longer(cols = c(CO2_umolL, CH4_umolL)) %>%
  filter(value >= 0) %>%
  mutate(Ecosystem = "Streams and rivers")
```

## Step 4: Oceans

```{r oceans}
#https://memento.geomar.de/contact.html
ocean <- readxl::read_excel("../CompiledData/nbparameter_data_export_2026-07-27.xls") %>%
  mutate(ch4 = as.numeric(ch4))

ocean_sum <- ocean %>%
  mutate(name = "CH4_umolL",
         value = ch4 / 1000) %>% # convert nmol/umol
  filter(!ch4 == -999,
         ch4 >= 0,
         !grepl("estuary", Label, ignore.case = T)) %>%
  select(name, value) %>%
  mutate(Ecosystem = "Oceans")

ocean_surface <- ocean %>%
  mutate(name = "CH4_umolL",
         value = ch4 / 1000) %>% # convert nmol/umol
  filter(!ch4 == -999,
         ch4 >= 0,
         !grepl("estuary", Label, ignore.case = T),
         `Depth [m]`<5) %>%
  mutate(Ecosystem = "Ocean\nsurface")

ocean_surface %>%
  ggplot(aes(x = Latitude, y=Longitude))+
  geom_point()
```

## Step 5: Estuaries

```{r estuaries}
#https://datadryad.org/dataset/doi:10.5061/dryad.7m0cfxpwz
#https://onlinelibrary.wiley.com/doi/abs/10.1111/gcb.16233?casa_token=mjCDhe9Wxa4AAAAA:Y_SAwfde15RDOpb_M8HZE4l-mftS2RXxFXwNPFJKI1aCas4UiiWbbw1Cdj-Urv3Pxcxg8lv7Y0Ym08Y
estuary <- read_csv("../CompiledData/estuary.csv")

estuary_sum <- estuary %>%
  select(CH4_umolL_Mean) %>%
  rename(CH4_umolL = CH4_umolL_Mean) %>%
  filter(!is.na(CH4_umolL),
         CH4_umolL >= 0) %>%
  pivot_longer(CH4_umolL) %>%
  mutate(Ecosystem = "Estuaries")
```

## Step 6: Wetlands

```{r wetlands}
#https://data.usgs.gov/datacatalog/data/USGS:59a86e39e4b0421949a84627
#https://www.frdr-dfdr.ca/repo/dataset/31a0f83a-f1f9-4517-8a6b-6d0674eee87b
#https://portal.edirepository.org/nis/mapbrowse?packageid=edi.1601.1
#https://borealisdata.ca/dataset.xhtml?persistentId=doi:10.5683/SP3/CJH4YV

logozzo <- readxl::read_excel(here::here("CompiledData","Logozzo_ppr_ghg_2021.xlsx"),
                              sheet = 2)
wetland_logozzo <- logozzo %>%
  mutate(
    T_K = as.numeric(Water.Temp) + 273.15,
    K_CO2 = exp(
      -58.0931 +
      90.5069 * (100 / T_K) +
      22.2940 * log(T_K / 100)
    ),
    K_CH4 = exp(
      -68.8862 +
      101.4956 * (100 / T_K) +
      28.7314 * log(T_K / 100)
    ),
    CO2_umolL = as.numeric(pCO2.insitu) * K_CO2,
    CH4_umolL = as.numeric(pCH4.insitu) * K_CH4
  ) %>%
  group_by(Site_name) %>%
  summarize(CH4_umolL = mean(CH4_umolL, na.rm = T),
            CO2_umolL = mean(CO2_umolL, na.rm = T)) %>%
  pivot_longer(cols = c(CO2_umolL, CH4_umolL)) %>%
  mutate(Ecosystem = "Freshwater wetlands")

lloreda <- read_csv(here::here("CompiledData","Merged_synoptic.csv")) %>%
  filter(Sample_Type == "SW",
         !is.na(Site_new), # These aren't in the paper
         !Site_ID == "ND-outlet"
         ) 

wetland_lloreda <- lloreda %>%
  group_by(Site_ID) %>%
  summarize(CH4_umolL = mean(CH4_uM, na.rm = T),
         CO2_umolL = mean(CO2_uM, na.rm = T)) %>%
  pivot_longer(cols = c(CO2_umolL, CH4_umolL)) %>%
  mutate(Ecosystem = "Freshwater wetlands")

wetland_bansal <- read_csv(here::here("CompiledData","dissolved_gas.csv")) %>%
  mutate(across(where(is.numeric), ~na_if(.x, -9999))) %>%
  mutate(
    T_K = S_W_TEMP_C + 273.15,
    K_CO2 = exp(
      -58.0931 +
      90.5069 * (100 / T_K) +
      22.2940 * log(T_K / 100)
    ),
    K_CH4 = exp(
      -68.8862 +
      101.4956 * (100 / T_K) +
      28.7314 * log(T_K / 100)
    ),
    CO2_umolL = as.numeric(CO2_CONC) * K_CO2,
    CH4_umolL = as.numeric(CH4_CONC) * K_CH4
  ) %>%
  group_by(WETLAND_ID) %>%
  summarize(CH4_umolL = mean(CH4_umolL, na.rm = T),
         CO2_umolL = mean(CO2_umolL, na.rm = T)) %>%
  pivot_longer(cols = c(CO2_umolL, CH4_umolL)) %>%
  mutate(Ecosystem = "Freshwater wetlands")

wetland_hassan <- read_csv("../CompiledData/Inter-regional spatial data_peatland pools.csv") %>%
  group_by(`Peatland Pools`) %>%
  summarize(
    CO2_umolL = as.numeric(`CO2-C (mgL-1)`) * 1000 / 12.011,
    CH4_umolL = as.numeric(`CH4 (μgL-1)`) / 12.011 #confirmed this with figure in paper
  )%>%
  pivot_longer(cols = c(CO2_umolL, CH4_umolL)) %>%
  mutate(Ecosystem = "Freshwater wetlands")

wetland <- bind_rows(
  wetland_bansal,
  wetland_lloreda,
  wetland_logozzo,
  wetland_hassan
)
```

## Step 7: Combine and plot

```{r combine_and_plot}

factor_levels = c("Oceans", "Estuaries", "Streams and rivers", "Freshwater wetlands",
                  "Lake\nsurface", "Lake\nbottom")

comb <- bind_rows(stream_sum,
                  tropical_rivers_sum,
                  estuary_sum,
                  ocean_sum,
                  wetland,
                  #ocean_surface,
                  flow) %>%
  bind_rows(layer_means) %>%
  mutate(Ecosystem = factor(Ecosystem,
                            factor_levels)) %>%
  mutate(name = recode_values(name,
                              "CH4_umolL"~"CH[4]",
                              "CO2_umolL"~"CO[2]"),
         study = ifelse(grepl("Lake", Ecosystem),
                        "This study",
                        NA)) %>%
  filter(value >= 0) %>%
  ungroup() 

text_init = comb %>%
  group_by(Ecosystem, name, study) %>%
  summarize(n = sum(!is.na(value)),
            median = median(value, na.rm = T),
            label_median = paste0("median =\n", format(median, drop0trailing = T, digits = 1)),
            label = paste0("n =\n", n))

# Consistent text positions
text <- comb %>%
  group_by(name) %>%
    mutate(
      value = if_else(
        value == 0,
        min(value[value > 0], na.rm = TRUE),
        value,
        missing = NA
      )
    )%>%
  summarize(
    xmin = min(value, na.rm = TRUE),
    xmax = max(value, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    xpos = 10^(log10(xmin) + 0.01 * (log10(xmax) - log10(xmin))),
    xpos2 = 10^(log10(xmax) - 0.01 * (log10(xmax) - log10(xmin))),
    ypos = 10^(log10(xmax) + 0.14 * (log10(xmax) - log10(xmin)))
  ) %>%
  left_join(text_init, by = "name") %>%
  mutate(
    #xpos = ifelse(grepl("Ocean", Ecosystem) & name == "CH[4]",
         #              xpos2, xpos),
         #hjust = ifelse(grepl("Ocean", Ecosystem) & name == "CH[4]",
         #              1, 0),
    xpos = xpos2,
         hjust = 1) %>%
  mutate(label = ifelse((Ecosystem == "Oceans"&name == "CH[4]") |
                          (Ecosystem == "Streams and rivers"&name == "CO[2]"), 
                        label, sub("n =\n", "", label)),
         label_median = ifelse((Ecosystem == "Oceans"&name == "CH[4]") |
                          (Ecosystem == "Streams and rivers"&name == "CO[2]"),
                        label_median, sub("median =\n", "", label_median)))

row_height <- 0.9 #cm

n_ch4 <- n_distinct(filter(comb, name == "CH[4]")$Ecosystem)
n_co2 <- n_distinct(filter(comb, name == "CO[2]")$Ecosystem)

final2 <- make_ecosystem_plot_wide(comb, text)
final2
ggsave(plot = final2, "../Figures/ecosystem_comparison2.png", 
       width = 5.5, height = 3.5)
```

## Step 8: Statistics

```{r stats}
library(dplyr)
library(rstatix)

kw_results <- comb %>%
  filter(value >= 0) %>%
  group_by(name) %>%
  rstatix::kruskal_test(value ~ Ecosystem)

kw_results

pairwise <- comb %>%
  filter(value >= 0) %>%
  mutate(value = log10(value)) %>%
  group_by(name) %>%
  rstatix::dunn_test(value ~ Ecosystem)

attributes(pairwise)

pairwise

comb %>%
  count(name, Ecosystem, value) %>%
  group_by(name, Ecosystem) %>%
  filter(name == "CO[2]") %>%
  arrange(desc(n), .by_group = F)
```
